---
title: "Tidy Tuesday #3"
author: "PUT YOUR NAME HERE"
format: pdf
---

## Assignment Description

This is the third tidy tuesday assignment! This is another choose your own dataset assignment because the datasets haven't been very good for exploration this year so far. 

As usual, I want you to do three separate tasks. One should be a plot (ggplot), one should involve creating a summary or summaries (dplyr), and one should involve working with text (stringr). Your analysis should include a discussion of what you are trying to do and what it tells you about the data. 

To choose a dataset, do the following:

-   Go to <https://github.com/rfordatascience/tidytuesday>
-   Choose a dataset from the six years of available data. When you find a dataset that you want, make sure to click on the hyperlink in the data column of the table.
-   Find "The Data" subsection of the page and copy the "Read directly from GitHub" code.
-   Place that code in the chunk below labeled "load the dataset".



```{r load packages}
#| echo: false
#| message: false

library("tidyverse");theme_set(theme_bw())
```

```{r load the dataset}

```

```{r analysis 1}

```

Discussion:

```{r analysis 2}

```

Discussion:

```{r analysis 3}

```

Discussion:
