---
title: "Tidy Tuesday #1"
author: "PUT YOUR NAME HERE"
format: pdf
---

## Assignment Description

This is your first Tidy Tuesday data analysis assignment! As I mentioned before, Tidy Tuesday is a collective that releases a dataset each week for people to use to strength their data science skills. That's what we are going to do here!

Right now, the main thing you know how to do is make graphs, so that's what we are going to do here! For this assignment, I want you to explore the a dataset of your choice by making at least three different plots. At least one of the plots need to contain two or more different variables. In addition to the plot, you need to write a discussion about why you chose the specific plot type and variables as well as what that plot tells you about the data.

To choose a dataset, do the following:

-   Go to <https://github.com/rfordatascience/tidytuesday>
-   Choose a dataset from the six years of available data. When you find a dataset that you want, make sure to click on the hyperlink in the data column of the table.
-   Find "The Data" subsection of the page and copy the "Read directly from GitHub" code.
-   Place that code in the chunk below labeled "load the dataset".



```{r load packages}
#| echo: false
#| message: false

library("tidyverse");theme_set(theme_bw())
```

```{r load the dataset}

```

```{r analysis 1}

```

Discussion:

```{r analysis 2}

```

Discussion:

```{r analysis 3}

```

Discussion:
